---
title: "Project 2: Exploratory Analysis of NOAA Storm Database"
author: "gspinnadauhu"
date: "March 3, 2017"
output: html_document
---

#### Synopsis


#### Introduction
Storms and other severe weather events can cause both public health and economic problems for communities and municipalities. Many severe events can result in fatalities, injuries, and property damage, and preventing such outcomes to the extent possible is a key concern.

This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

#### Data
The data for this assignment come in the form of a comma-separated-value file compressed via the bzip2 algorithm to reduce its size.

#####Sources:
[Storm Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)
[Database Documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf)
[FAQ](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2FNCDC%20Storm%20Events-FAQ%20Page.pdf)

The events in the database start in the year 1950 and end in November 2011. In the earlier years of the database there are generally fewer events recorded, most likely due to a lack of good records. More recent years should be considered more complete.

#### Data Processing

Decompress and load the file as a dataframe.
```{r}
stormdata<-read.table("./data/repdata%2Fdata%2FStormData.csv.bz2",header=TRUE,sep=",")
```
Taking a look at the structure of the dataframe.
```{r}
str(stormdata)
```
Most variables were loaded correctly, but some need to be reclassed as dates or factors.
Using dplyr + mutate_each to reclass
```{r}
library("dplyr")
library("lubridate")
stormdata<-stormdata %>%
  mutate_each(funs(factor),c(STATE__,COUNTY)) %>%
  mutate_each(funs(mdy_hms),c(BGN_DATE,END_DATE))
```

#### Results



