---
title: "Project 2: Exploratory Analysis of NOAA Storm Database"
author: "gspinnadauhu"
date: "March 3, 2017"
output: html_document
---

#### Synopsis


#### Introduction
Storms and other severe weather events can cause both public health and economic problems for communities and municipalities. Many severe events can result in fatalities, injuries, and property damage, and preventing such outcomes to the extent possible is a key concern.

This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

#### Data
The data for this assignment come in the form of a comma-separated-value file compressed via the bzip2 algorithm to reduce its size.

#####Sources:
[Storm Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)
[Database Documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf)
[FAQ](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2FNCDC%20Storm%20Events-FAQ%20Page.pdf)

The events in the database start in the year 1950 and end in November 2011. In the earlier years of the database there are generally fewer events recorded, most likely due to a lack of good records. More recent years should be considered more complete.

#### Data Processing
##### Loading & Formatting the data
Decompress and load the file as a dataframe.
```{r}
data<-read.table("./data/repdata%2Fdata%2FStormData.csv.bz2",header=TRUE,sep=",")
```
Taking a look at the structure of the dataframe.
```{r}
str(data)
```
Selecting only the pertinent variables for this asignment: EVTYPE, FATALITIES, INJURIES, PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP, REFNUM. These are put into a new dataframe called stormdata.
```{r}
library(dplyr)
stormdata<-select(data,EVTYPE,FATALITIES:CROPDMGEXP,REFNUM)
```

##### Questions
1.  Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?
    The variables describing population health effects are FATALITIES and INJURIES. Using dplyr summarise, in order to summarize these variables by EVTYPE. These summaries are then saved in dataframes called sum_fatalities and sum_injuries. Renaming quantitative variables to Total_F and Total_I
    ```{r}
    library(dplyr)
    sum_fatalities<-summarise(group_by(stormdata,EVTYPE),sum(FATALITIES))
    sum_injuries<-summarise(group_by(stormdata,EVTYPE),sum(INJURIES))
    names(sum_fatalities)[[2]]<-"Total_F"
    names(sum_injuries)[[2]]<-"Total_I"
    ```
    Now filtering each dataframe for the top 10 largest number of fatalities & injuries.
    Assigning these to new dataframes called fatalities_10 & injuries_10 and putting in descending order of totals.
    ```{r}
    library(dplyr)
    fatalities_10<-top_n(sum_fatalities,10,Total_F)
    injuries_10<-top_n(sum_injuries,10,Total_I)
    fatalities_10<-arrange(fatalities_10,desc(Total_F))
    injuries_10<-arrange(injuries_10,desc(Total_I))
    ```
2.  Across the United States, which types of events have the greatest economic consequences?

#### Results



